```{r}
library(tidyverse)
library(shiny)
```
```{r}
df <- read.csv('Data/albums.csv', stringsAsFactors = F)
df <- df %>% 
  filter(Number.of.reviews >= 1, !str_detect(Band, '\\|')) # set a minimum of 1 review; no splits (splits are when Band = band_a | band_b)
```
```{r}
# split the genres by early and later
df <- df %>% mutate(genre_early = trimws(str_replace_all(Genre,'\\(early.*', '')),
                    genre_early = case_when(str_sub(genre_early, start=-5) == '(mid)' ~ str_extract(genre_early, '.+?(?=,)'),
                                            TRUE ~ genre_early),
              genre_later = case_when(str_sub(Genre, start=-5) == '(mid)' ~ genre_early,
                                      # if (early, later), (early; later), (early & later), (early/later) is found
                                      str_detect(Genre, '\\(early(,|;|\\&) later\\)') ~ genre_early,
                                      # between (mid) and (later)
                                      str_detect(Genre, '\\(mid\\)') ~ str_extract(Genre, '(?<=\\(mid\\)?, ).+?(?= \\(later\\))'),
                                      #between ', ' and '(later)'
                                      str_detect(Genre, '(later)') ~ str_extract(Genre, '(?<=, ).+?(?= \\(later\\))'), 
                                      TRUE ~ Genre)
              )
```

```{r}
# get the root genres from the early, late genre columns
df <- df %>% mutate(genre_early_secondary = case_when(# in between '/' and 'Metal'
                                           # str_detect(genre_early, '/') ~ str_extract(genre_early, '(?<=[:punct:])[:alpha:](?=Metal)'),
                                           # get the word before 'Metal'
                                           str_detect(genre_early, 'Metalcore') ~ 'Metalcore',
                                           str_detect(genre_early, 'Metal') ~ str_extract(genre_early, '\\w+(?=(\\-|\\s)+Metal)'),
                                           str_detect(genre_early, 'NWOBHM') ~ 'NWOBHM',
                                           str_detect(genre_early, 'Rock') ~ 'Rock',
                                           # get the word containing -core
                                           str_detect(genre_early, 'core') ~ str_extract(genre_early, '(\\w+core)'),
                                           TRUE ~ genre_early),
              genre_later_secondary = case_when(# in between '/' and 'Metal'
                                           # str_detect(genre_later, '/') ~ str_extract(genre_later, '(?<=[:punct:])[:alpha:](?=Metal)'),
                                           str_detect(genre_early, 'Metalcore') ~ 'Metalcore',
                                           str_detect(genre_later, 'Metal') ~ str_extract(genre_later, '\\w+(?=(\\-|\\s)+Metal)'),
                                           str_detect(genre_later, 'NWOBHM') ~ 'NWOBHM',
                                           str_detect(genre_later, 'Rock') ~ 'Rock',
                                           str_detect(genre_later, 'core') ~ str_extract(genre_later, '(\\w+core)'),
                                           TRUE ~ genre_later)
              )# %>% filter(!str_detect(genre_early, 'Metal')) %>% distinct(genre_later_main, .keep_all=T) %>% select(genre_later, genre_later_main)
```
```{r}
df <- df %>% mutate(genre_early_main = case_when(str_detect(genre_early_secondary, 'Thrash|Speed|Crosso') ~ 'Thrash Metal',
                                           str_detect(genre_early_secondary, 'NWOBHM|Heavy|Trad') ~ 'Heavy Metal',
                                           # exclude deathcore
                                           (str_detect(genre_early_secondary, 'Death|Extreme') & !str_detect(genre_early_secondary, 'core')) ~ 'Death Metal',
                                           str_detect(genre_early_secondary, 'Ambient|garde|Atmos|Experi|Drone|Electro|Post') ~ 'Ambient',
                                           str_detect(genre_early_secondary, 'Folk|Viking|Pagan|Pirate|Medieval|Celtic') ~ 'Folk Metal',
                                           str_detect(genre_early_secondary, 'Black|Dark') ~ 'Black Metal',
                                           str_detect(genre_early_secondary, 'Gothic|Indust|Groove|Nu|South|New') ~ 'Nu Metal',
                                           str_detect(genre_early_secondary, 'Power|Neocla|Melod') ~ 'Power Metal',
                                           str_detect(genre_early_secondary, 'Doom|Sludge|Stoner') ~ 'Doom Metal',
                                           str_detect(genre_early_secondary, 'Prog|Symph|Oper|Classi') ~ 'Progressive Metal',
                                           str_detect(genre_early_secondary, 'Rock|AOR|Alter|Glam|Punk|Grunge') ~ 'Rock',
                                           str_detect(genre_early_secondary, 'core|Crust|grind') ~ 'Metalcore',
                                           TRUE ~ 'Other'),
              genre_later_main = case_when(str_detect(genre_later_secondary, 'Thrash|Speed|Crosso') ~ 'Thrash Metal',
                                             str_detect(genre_later_secondary, 'NWOBHM|Heavy|Trad') ~ 'Heavy Metal',
                                             # exclude deathcore
                                             (str_detect(genre_later_secondary, 'Death|Extreme') & !str_detect(genre_later_secondary, 'core')) ~ 'Death Metal',
                                             str_detect(genre_later_secondary, 'Ambient|garde|Atmos|Experi|Drone|Electro|Post') ~ 'Ambient',
                                             str_detect(genre_later_secondary, 'Folk|Viking|Pagan|Pirate|Medieval|Celtic') ~ 'Folk Metal',
                                             str_detect(genre_later_secondary, 'Black|Dark') ~ 'Black Metal',
                                             str_detect(genre_later_secondary, 'Gothic|Indust|Groove|Nu|South|New') ~ 'Nu Metal',
                                             str_detect(genre_later_secondary, 'Power|Neocla|Melod') ~ 'Power Metal',
                                             str_detect(genre_later_secondary, 'Doom|Sludge|Stoner') ~ 'Doom Metal',
                                             str_detect(genre_later_secondary, 'Prog|Symph|Oper|Classi') ~ 'Progressive Metal',
                                             str_detect(genre_later_secondary, 'Rock|AOR|Alter|Glam|Punk|Grunge') ~ 'Rock',
                                             str_detect(genre_later_secondary, 'core|Crust|grind') ~ 'Metalcore',
                                             TRUE ~ 'Other')
)# %>% filter(genre_later_main == 'Other') %>% distinct(genre_early_secondary, .keep_all=T) %>% select(Band, genre_early,genre_early_secondary,genre_early_main)
```


```{r}
# get the earliest location of a band
df <- df %>% mutate(location_early = trimws(str_replace_all(Location,'\\(early.*', '')), # keep everything before '(early)'
              location_early = case_when(str_sub(location_early, start=-5) == '(mid)' ~ str_extract(location_early, '.+?(?=,)'),
                                         str_detect(location_early, '\\/') ~ str_replace_all(location_early, '\\/.*', ''),
                                         TRUE ~ location_early)
)
```

```{r}
df <- df %>% 
  #select(genre_early_main, genre_early) %>% 
  #distinct(genre_early, .keep_all=T) %>% 
  mutate(genre_early_stripped = trimws(str_remove_all(str_replace_all(str_remove(genre_early, genre_early_main), '\\/|,', ' '), 'Metal')),
         genre_later_stripped = trimws(str_remove_all(str_replace_all(str_remove(genre_later, genre_later_main), '\\/|,', ' '), 'Metal')))# %>%distinct(genre_early_stripped, .keep_all=T)
```
```{r}
```
df <- select(-genre_tags)
df
```{r}
df %>% select(-genre_tags) %>% write_csv('albums_cleaned.csv')
vapply(df$genre_tags, paste, collapse = ", ")
```



```{r}
# narrow down genres
df %>% distinct(genre_later_secondary, .keep_all=T) %>% select(Band, Genre, genre_early_secondary, genre_later_secondary)
```
```{r}
df %>% 
  group_by(genre_early_secondary) %>%
  summarise(no_rows = length(genre_early_secondary)) %>% arrange(desc(no_rows))
```

```{r}
df <- df %>% mutate(genre_early_main = case_when(str_detect(genre_early_secondary, 'Thrash|Speed|Crosso') ~ 'Thrash Metal',
                                           str_detect(genre_early_secondary, 'NWOBHM|Heavy|Trad') ~ 'Heavy Metal',
                                           # exclude deathcore
                                           (str_detect(genre_early_secondary, 'Death|Extreme') & !str_detect(genre_early_secondary, 'core')) ~ 'Death Metal',
                                           str_detect(genre_early_secondary, 'Ambient|garde|Atmos|Experi|Drone|Electro|Post') ~ 'Ambient',
                                           str_detect(genre_early_secondary, 'Folk|Viking|Pagan|Pirate|Medieval|Celtic') ~ 'Folk Metal',
                                           str_detect(genre_early_secondary, 'Black|Dark') ~ 'Black Metal',
                                           str_detect(genre_early_secondary, 'Gothic|Indust|Groove|Nu|South|New') ~ 'Nu Metal',
                                           str_detect(genre_early_secondary, 'Power|Neocla|Melod') ~ 'Power Metal',
                                           str_detect(genre_early_secondary, 'Doom|Sludge|Stoner') ~ 'Doom Metal',
                                           str_detect(genre_early_secondary, 'Prog|Symph|Oper|Classi') ~ 'Progressive Metal',
                                           str_detect(genre_early_secondary, 'Rock|AOR|Alter|Glam|Punk|Grunge') ~ 'Rock',
                                           str_detect(genre_early_secondary, 'core|Crust|grind') ~ 'Metalcore',
                                           TRUE ~ 'Other'),
              genre_later_main = case_when(str_detect(genre_later_secondary, 'Thrash|Speed|Crosso') ~ 'Thrash Metal',
                                             str_detect(genre_later_secondary, 'NWOBHM|Heavy|Trad') ~ 'Heavy Metal',
                                             # exclude deathcore
                                             (str_detect(genre_later_secondary, 'Death|Extreme') & !str_detect(genre_later_secondary, 'core')) ~ 'Death Metal',
                                             str_detect(genre_later_secondary, 'Ambient|garde|Atmos|Experi|Drone|Electro|Post') ~ 'Ambient',
                                             str_detect(genre_later_secondary, 'Folk|Viking|Pagan|Pirate|Medieval|Celtic') ~ 'Folk Metal',
                                             str_detect(genre_later_secondary, 'Black|Dark') ~ 'Black Metal',
                                             str_detect(genre_later_secondary, 'Gothic|Indust|Groove|Nu|South|New') ~ 'Nu Metal',
                                             str_detect(genre_later_secondary, 'Power|Neocla|Melod') ~ 'Power Metal',
                                             str_detect(genre_later_secondary, 'Doom|Sludge|Stoner') ~ 'Doom Metal',
                                             str_detect(genre_later_secondary, 'Prog|Symph|Oper|Classi') ~ 'Progressive Metal',
                                             str_detect(genre_later_secondary, 'Rock|AOR|Alter|Glam|Punk|Grunge') ~ 'Rock',
                                             str_detect(genre_later_secondary, 'core|Crust|grind') ~ 'Metalcore',
                                             TRUE ~ 'Other')
)# %>% filter(genre_later_main == 'Other') %>% distinct(genre_early_secondary, .keep_all=T) %>% select(Band, genre_early,genre_early_secondary,genre_early_main)
```
```{r}
df <- df %>% mutate(location_early = trimws(str_replace_all(Location,'\\(early.*', '')), # keep everything before '(early)'
              location_early = case_when(str_sub(location_early, start=-5) == '(mid)' ~ str_extract(location_early, '.+?(?=,)'),
                                         str_detect(location_early, '\\/') ~ str_replace_all(location_early, '\\/.*', ''),
                                         TRUE ~ location_early)
)
df %>% distinct(genre_early_secondary)
df %>% distinct(genre_early)
# get genre tags
df <- df %>% 
  select(genre_early_main, genre_early) %>% 
  distinct(genre_early, .keep_all=T) %>% 
  mutate(genre_early_stripped = trimws(str_remove_all(str_replace_all(str_remove(genre_early, genre_early_main), '\\/|,', ' '), 'Metal')),
         genre_tags = str_split(genre_early_stripped, '. ')) %>%
  distinct(genre_early_stripped, .keep_all=T)
```

```{r}
# todo: clean later
# get everything AFTER the last instance of '), ' and BEFORE (current) or (present) or (now)
df %>% mutate(location_early = trimws(str_replace_all(Location,'\\(early.*', '')), # keep everything before '(early)'
              location_early = case_when(str_sub(location_early, start=-5) == '(mid)' ~ str_extract(location_early, '.+?(?=,)'),
                                         str_detect(location_early, '\\/') ~ str_replace_all(location_early, '\\/.*', ''),
                                         TRUE ~ location_early),
              #location_later = str_replace_all(Location, location_early, ''),
              location_later = case_when(
                        #str_detect(Location, '\\(now\\)') ~ str_extract(Location, '(?<=\\), ).+?(?= \\(now\\))'),
                        #match all chars until '), ', anchor to end of string with $ #match all chars until '), ', anchor to end of string with $
                        #str_detect(Location, '\\(now\\)') ~ str_extract(Location, '(\\),.*)$'),
                        #str_detect(Location, '\\(now\\)') ~ str_extract(Location, '\\), (?!\\), ).*'), 
                        str_sub(Location, start=-5) == '(mid)' ~ location_early, # keep
                        # if (early, later), (early; later), (early & later), (early/later) is found
                        str_detect(Location, '\\(early(,|;|\\&) later\\)') ~ location_early, # keep
                        # between (mid) and (later)
                        str_detect(Location, '\\(mid\\)') ~ str_extract(Location, '(?<=\\(mid\\)?, ).+?(?= \\(later\\))'),
                        str_detect(Location, '\\/') ~ str_extract(Location, '(?<=\\/).*$'),
                        # between ', ' and '(later)', get 
                        str_detect(Location, '(later)') ~ str_extract(Location, '(?<=, ).+?(?= \\(later\\))'), 
                        TRUE ~ Location)
              ) %>% filter(location_early!=location_later) %>% distinct(Location, location_early, location_later)

df %>% filter(str_detect(Location, '\\(now|current|present\\)')) %>% select(Band, Location) %>% distinct(Location, .keep_all=T)
```
```{r}

```
# to-do:
# - clean locations
# - list genre tags
# get the tags from a genre; are they melodic, symphonic, brutal etc.?


```{r}

split_genres <- function(x){
  unlist(str_split(x, '/|,|\\|'))
}
df %>% mutate(more_genres = str_split(trimws(Genre), '\\((early|mid|later)\\),')) 
df$more_genres[1:10]
df %>% mutate(more_genres = split_genres(more_genres))
df$genre_split <- df$more_genres %>% lapply(split_genres)
df %>% distinct(Band, .keep_all=T) %>% filter(Number.of.reviews > 10) %>% select(Band, Genre, genre_split)
```

```{r}
df %>% arrange(Release.date)
```

```{r}
df %>% filter(Number.of.reviews > 10) %>% arrange(desc(Average.rating))
```
```{r}
library(datetime)
df <- df %>% mutate(Release.date = as.date(Release.date)) %>% filter(!is.na(Release.date))
```
```{r}
unique(df$Genre)[1:50]
```

To do:
- encapsulate genres into select few (Drone, Sludge, Doom -> Doom;)
- add "Metal" if genre doesn't include metal
- clean locations

```{r}
df <- df %>% 
    mutate(Genres = str_split(
      str_replace_all(Genre, 
                      '\\((early|mid|later)\\)', 
                      ''
                      ), 
      '/|,|\\|'
      )
      )

#df %>% select(Band, Genre, Genres) %>% distinct(Genre)
  # mutate(Genre = case_when(!(str_detect(Genre, 'Metal')) ~ paste0(Genre, ' Metal'),
  #                          TRUE ~ Genre))

#df %>% mutate(Genre_main = case_when())
```
```{r}
df %>% 
  filter(Number.of.reviews > 8) %>%
  select(Band, Genre, Genres) %>% 
  filter(str_detect(Genre,'(later)')) %>% 
  distinct(Band, .keep_all=T)
```
```{r}
(df$Genre %>% unique())[1:50]
```
```{r}
df %>% pull(Genre) %>% unique()
```
```{r}
df %>% filter(str_detect(Genre, '(later)')) %>% head(150) %>% pull(Genre) %>% unique() 
```

```{r}
# to-do: put genres into early, late columns

#df <- df %>% filter(Number.of.reviews >= 1, !str_detect(Band, '\\|'))
df <- df %>% mutate(genre_early = trimws(str_replace_all(Genre,'\\(early.*', '')),
                    genre_early = case_when(str_sub(genre_early, start=-5) == '(mid)' ~ str_extract(genre_early, '.+?(?=,)'),
                                            TRUE ~ genre_early),
              genre_later = case_when(str_sub(Genre, start=-5) == '(mid)' ~ genre_early,
                                      # if (early, later), (early; later), (early & later), (early/later) is found
                                      str_detect(Genre, '\\(early(,|;|\\&) later\\)') ~ genre_early,
                                      # between (mid) and (later)
                                      str_detect(Genre, '\\(mid\\)') ~ str_extract(Genre, '(?<=\\(mid\\)?, ).+?(?= \\(later\\))'),
                                      #between ', ' and '(later)'
                                      str_detect(Genre, '(later)') ~ str_extract(Genre, '(?<=, ).+?(?= \\(later\\))'), 
                                      TRUE ~ Genre)
              )
# df %>% distinct(Band, .keep_all=T) %>% filter(str_detect(Genre, '\\(mid\\)'), !str_detect(Band, '\\|')) %>% head(50) %>% select(Band, Genre, genre_early, genre_later)
df %>% distinct(Band, .keep_all=T) %>% select(Band, Genre, genre_early, genre_later)# %>% tail(100)
```
```{r}
df %>% mutate(genre_early_main = case_when(# in between '/' and 'Metal'
                                           # str_detect(genre_early, '/') ~ str_extract(genre_early, '(?<=[:punct:])[:alpha:](?=Metal)'),
                                           # get the word before 'Metal'
                                           str_detect(genre_early, 'Metal') ~ str_extract(genre_early, '\\w+(?=(\\-|\\s)+Metal)'),
                                           str_detect(genre_early, 'NWOBHM') ~ 'NWOBHM',
                                           str_detect(genre_early, 'Rock') ~ 'Rock',
                                           str_detect(genre_early, 'core') ~ str_extract(genre_early, '(\\w+core)'),
                                           TRUE ~ genre_early),
              genre_later_main = case_when(# in between '/' and 'Metal'
                                           # str_detect(genre_later, '/') ~ str_extract(genre_later, '(?<=[:punct:])[:alpha:](?=Metal)'),
                                           # get the word before 'Metal'
                                           str_detect(genre_later, 'Metal') ~ str_extract(genre_later, '\\w+(?=(\\-|\\s)+Metal)'),
                                           str_detect(genre_later, 'NWOBHM') ~ 'NWOBHM',
                                           str_detect(genre_later, 'Rock') ~ 'Rock',
                                           str_detect(genre_later, 'core') ~ str_extract(genre_later, '(\\w+core)'),
                                           TRUE ~ genre_later)
              ) %>% 
  
  #filter(!str_detect(genre_early, 'Metal')) %>% 
  distinct(genre_later_main, .keep_all=T) %>% 
  select(genre_later, genre_later_main)
```

```{r}
nrow(df)
```


```{r}
df %>% filter(str_detect(Genre, '(mid)'), Number.of.reviews>5)
```


```{r}
split_genres <- function(x){
  unlist(str_split(x, '/|,|\\|'))
}
df <- df %>% mutate(more_genres = str_split(trimws(Genre), '\\((early|mid|later)\\),')) 
df$more_genres[1:10]
df %>% mutate(more_genres = split_genres(more_genres))
df$genre_split <- df$more_genres %>% lapply(split_genres)
df %>% distinct(Band, .keep_all=T) %>% filter(Number.of.reviews > 10) %>% select(Band, Genre, genre_split)
#(df$more_genres %>% unique() %>% lapply(split_genres))[1:50]
#df$more_genres %>% unique() %>% lapply(str_split, '/') %>% lapply(unlist)
```

```{r}
df %>% distinct(Band, .keep_all=T) %>% filter(Number.of.reviews > 10)
```
```{r}
df %>% distinct(Band, .keep_all=T) %>% filter(Number.of.reviews > 5) %>% select(Band, Genre)
```

```{r}
df %>% distinct(Band, .keep_all=T) %>% filter(Number.of.reviews > 10) %>% pull(more_genres) %>% unique()
```


```{r}
df %>% pull(Genres) %>% unique()
```


```{r}
df %>% 
  filter(Number.of.reviews > 5) %>% 
  # split at the backslash i.e. black/thrash metal -> 'black', 'thrash metal'
  mutate(Genre = trimws(str_replace(Genre, '([:punct:]).*', ''))) %>%
  mutate(Genre = case_when(!(str_detect(Genre, 'Metal')) ~ paste0(Genre, ' Metal'),
                           TRUE ~ Genre)) %>% # %>% pull(Genre) %>% unique()
  mutate(Genre_main = case_when(str_detect(Genre, 'Dark|Black') ~ 'Black Metal',
                                str_detect(Genre, 'Heavy|Traditional|NWOBHM') ~ 'Heavy Metal',
                                str_detect(Genre, 'Doom|Sludge|Stoner|Drone') ~ 'Doom Metal',
                                str_detect(Genre, 'Speed|Thrash|Crossover') ~ 'Thrash Metal',
                                str_detect(Genre, 'Groove|Industrial|Southern') ~ 'Groove Metal',
                                str_detect(Genre, 'Brutal|Technical|Death') ~ 'Death Metal',
                                str_detect(Genre, 'Power|Symphonic|Progressive') ~ 'Power Metal',
                                str_detect(Genre, 'Pagan|Pirate|Folk|Viking') ~ 'Folk Metal',
                                str_detect(Genre, 'Crust|Core|grind|Grind') ~ 'Crust/Grindcore',
                                str_detect(Genre, 'Post|Experimental|Ambient|Avant') ~ 'Avant',
                                str_detect(Genre, 'Metalcore|Deathcore|Hardcore') ~ 'Metalcore',
                                str_detect(Genre, 'Rock') ~ 'Rock',
                                str_detect(Genre, 'Goth|Nu') ~ 'Nu Metal'
                                )
         ) %>% filter(is.na(Genre_main)) %>% distinct(Band, Genre, Genre_main)

#str_replace(unique(df$Genre)[1:25], '([:punct:]).*', '')
#str_subset(df$Genre, '(.*/)')
```
```{r}
df %>% filter(Number.of.reviews > 10) %>% mutate(Location = str_replace(Location, '\\(early\\).*', ''))
```

```{r}
df %>% arrange(desc(Number.of.reviews))

```
```{r}
df %>%
```

